---
output: github_document
layout: post
---

```{r setup, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
require(ggplot2)
require(stringr)
require(httr)
require(dplyr)
require(jsonlite)
require(tidyr)
library(gt)
library(tidyverse)
library(glue)

appraisal_history2019<-read.csv("D:\\Urban3\\Projects\\NC\\buncombe_co\\analyzed_data\\r_output\\Real_Estate_Appraisal_Tax_History_2019.csv", header=TRUE)
sales_data_table<-read.csv("D:\\Urban3\\Projects\\NC\\buncombe_co\\analyzed_data\\r_output\\sales_data_properties.csv", header=TRUE)
```

## Introduction

### Price Related Differential

IAAO definitation: Price-related differential. The mean (sales ratio) divided by the weighted mean (sales ratio). The statistic has a slight bias upward. Price-related differentials above 1.03 tend to indicate assessment regressivity; price-related differentials below 0.98 tend to indicate assessment progressivity

Fair & Equitable 2011: 

- "There is a bias in the measure such that PRD in the general range of 0.9 and 1.1 are inconclusive indicator of assessment regressivity (IAAO, 1980, 14)"
- "The 1990 Standard on Ratio Studies (IAAO 1990, 24-25) set forth a standard for PRD of 0.98 to 1.03"
- "The price-related differential (PRD) and weighted mean are sensitive to sales with high prices even if the ratios on higher priced sales do not appear unusual relative to other sales"

At best, PRD is 'an easy to calculate measure that provides "some notion of any association" between assessment ratios and property value. 

### Coefficient of Dispersion
Less than 15 for residential properties

## Calculating Indicators of Vertical Equity
```{r, eval=TRUE, message=FALSE, warning=FALSE, results=FALSE}
appraisal_history2019$totalvalue<-appraisal_history2019$LandVal+appraisal_history2019$BldgVal+appraisal_history2019$ImprVal ##Calculate total taxable value
appraisal_history2019_abbr<-appraisal_history2019 %>%
  distinct(ï..PIN,TaxYear, .keep_all = TRUE) %>% #The data set includes duplicated records. Remove duplicates.
  mutate(PIN=ï..PIN, #recode variable names
         year=TaxYear,
         landvalue=LandVal)%>%
  select(PIN, year, landvalue, totalvalue)
appraisal_history2019_abbr$PIN_YEAR=paste(appraisal_history2019_abbr$PIN, appraisal_history2019_abbr$year, sep="_")

#Extract the year from the data field
sales_year<-as.data.frame(sapply(strsplit(sales_data_table$SellDate, "-"), `[`,1))
names(sales_year)<-c("SALES_YEAR")
sales_data_table<-cbind(sales_data_table, sales_year)
sales_data_table$PIN<-str_replace_all(sales_data_table$PINN, "-", "")
sales_data_table$SALES_YEAR<-as.numeric(sales_data_table$SALES_YEAR)

#Extract the most recent sale record
sales_data_table_year<-sales_data_table %>% 
  filter(QualifiedSale=='Y')  %>% #Select qualified sales only
  filter(SALES_YEAR>2000)%>% #Filter sales only after 2000, since we only have appraisal values back to 2001
  filter(VacantLot=="False")%>%
  filter(SellingPrice>10000)%>%
  mutate(SALES_YEAR_1=(SALES_YEAR+1))%>% #Create a new date field that allows the sales to align with the following tax year
  select(PIN, SellingPrice, City, Subdivision, Class, SALES_YEAR, SALES_YEAR_1)
sales_data_table_year$PIN_YEAR=paste(sales_data_table_year$PIN, sales_data_table_year$SALES_YEAR_1, sep="_")

sales_appraised<-merge(appraisal_history2019_abbr, sales_data_table_year, by="PIN_YEAR")
sales_appraised$sales_ratio<-sales_appraised$totalvalue/sales_appraised$SellingPrice

sales_ratio<-sales_appraised%>%
  filter(Class=="RES")%>%
  group_by(SALES_YEAR)%>%
  mutate(selldecile = ntile(SellingPrice, 10),
         mean_ratio = mean(sales_ratio),
         median_ratio = median(sales_ratio),
         weighted_mean = sum(totalvalue)/sum(SellingPrice),
         ratio_std = sd(sales_ratio),
         index = 1,
         count = sum(index),
         median_diff=(sales_ratio-median_ratio),
         coeff_dispersion = (100/count)*(sum(abs(median_diff))/median_ratio),
         coeff_variation = ((100*ratio_std)/mean_ratio),
         price_rltd_diff = (mean_ratio/weighted_mean),
         Year = first(SALES_YEAR))

sales_ratio_summary<-sales_ratio%>%
  group_by(Year)%>%
  summarise(`Mean Ratio` = first(mean_ratio),
            `Meadian Ratio` = first(median_ratio),
            `Coeff. of Dispersion` = first(coeff_dispersion),
            `Coeff. of Variation` = first(coeff_variation),
            `Price Related Differential` = first(price_rltd_diff),
            `Transactions`=sum(index))
```

## Table 1
```{r, echo=FALSE, message=FALSE, warning=FALSE}
sales_ratio_summary %>%
  arrange(Year) %>%
  gt() %>%
  tab_header(
    title = "Indicators of Vertical Equity",
    subtitle = "Buncombe County, North Carolina")%>%
  fmt_number(columns = 2:6,
             decimals=2)
```
```
